BIVAN makefile, compiliation notes

### 2019-02-11 Paper.Rmd
changes to make ncsurrog throw an error when it should

For weak noise and using our nonlinear model (methods for Q2), copula statistics
were again very similar for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}).
Though there were again significant but small differences for some statistics
and simulations, the overall hypothesis that copula structure in the noise translates
to similar but not identical copula structure in output populations, at least
as measured with the tail dependence and correlation statistics we used, was
supported. Since many ecological models are nonlinear, this result provides the 
important expectation that there will be a Moran-effect-like correspondence 
between noise and model-output copula structure for typical ecological dynamics,
as long as the noise is small enough that dynamics stay relatively close
to the model equilibrium. Results that hold for "weak noise" in this sense
are common in ecology. 
We repeated the analysis with $r=X$, with similar results. The deterministic 
<!--***DAN: Shya, please replace X in the above line with the actual value(s)
of r that you used. But please be sure
the results were "similar". I encourage you to show me the actual results if there is 
any doubt in your mind.-->
one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
at $K$ when $r<1$ (undercompensating dynamics), but exhibits 
an oscillatory approach to $K$ when $r>1$ (overcompensating
dynamics).

For strong noise and using our nonlinear model (methods for Q2), 
copula statistics, generally speaking, were approximately similar 
between noise and model
outputs; however, similarity was reduced relative to the AR(1) 
and weak-noise cases, and, for a few simulations, tail dependence 
asymmetry statistics
even had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a C copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population model outputs; and likewise for SG and SJ copulas. Dynamics were
strongly nonlinear for the model parameters we used, and nevertheless similarities
between noise and model-output copula structure, according to the statistics 
we used, were more noticable than discrepancies.
We again repeated the analysis with $r=X$, with similar results.
<!--***DAN: Right, Shya? Please replace X with the actual value(s), but please be sure
the results were "similar". I encourage you to show me the actual results if there is 
any doubt in your mind.-->

@@

<!--We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.--> 
<!--
## Application : The Moran effect and copula structure in population

```{r read_RES_temp_aphid, echo=F}
RES_temp_aphid<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_winter_neg_temp.RDS")
```

Firstly, Our hypothesis was if a non-normal copula structure describes the dependence of an environmental
variable between two locations, then there can be a non-normal copula structure for the
dependence of populations in the locations that are influenced by the environmental variables. This assumption was inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations. In the previous section, we already tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment (having non-Normal copula structure) and found that populations were shaped by similar copula structure as that of local noises. Here, we investigated the probable cause of upper tail dependence in the leaf-curling plum aphid's firstflight date throughout all the locations. We assumed that average winter temperature which might be spatially synchronized and had asymmetric tail dependence cause the synchronized delay in leaf-plum curling aphid's appearence. We tested our hypothesis and indeed found the Moran effect on leaf-plum curling aphid's firstflight data induced by average winter temperature for the period of 35 years (1976-2010). We did pair wise model selection with negative average winter temperature (averaged over (n-1)$^{th}$ December to n$^{th}$ March temperature) from the selected 10 locations for which we did model selection with aphid-firstflight data. Negative winter temperature was considered for simulation to compare the aphid's firstflight results with temparature copula (aphid's firstflight date were negatively correlated with temperature or in other words positively correlated with negative temparature). Model selection results for negative average winter temperature in 10 $\times$ 10 matrices were represented in Table \ref{SM-tab_temp_aphid_bestcop} - Table \ref{SM-tab_temp_aphid_LTmUT}. In all cases `r min((RES_temp_aphid$gfc_numBS_success/RES_temp_aphid$gfc_numBS),na.rm=T)*100` $\%$ of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data. Table \ref{SM-tab_temp_aphid_LTmUT} showed `r sum(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw <0, na.rm=T)` out of `r (dim(RES_temp_aphid$gfc_p_CvM)[1]^2-dim(RES_temp_aphid$gfc_p_CvM)[1])-2*RES_temp_aphid$num_indep_loc_pair` location pair showed upper tail dependence in temperature copula, which can be considered as the Moran effect results causing dominant upper tail dependence in the copulae with aphid's firstflight date (Table \ref{SM-tab_ff_LTmUT}). -->

>>>>

For weak noise and using our nonlinear model (methods for Q2), copula statistics
were again very similar for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}).
Though there were again significant but small differences for some statistics
and simulations, the overall hypothesis that copula structure in the noise translates
to similar but not identical copula structure in output populations, at least
as measured with the tail dependence and correlation statistics we used, was
supported. Since many ecological models are nonlinear, this result provides the 
important expectation that there will be a Moran-effect-like correspondence 
between noise and model-output copula structure for typical ecological dynamics,
as long as the noise is small enough that dynamics stay relatively close
to the model equilibrium. Results that hold for "weak noise" in this sense
are common in ecology. 
We repeated the analysis with $r=1.3$, with similar results. The deterministic 
one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
at $K$ when $r<1$ (undercompensating dynamics), but exhibits 
an oscillatory approach to $K$ when $r>1$ (overcompensating
dynamics).

For strong noise and using our nonlinear model (methods for Q2), 
copula statistics, generally speaking, were approximately similar 
between noise and model
outputs; however, similarity was reduced relative to the AR(1) 
and weak-noise cases, and, for a few simulations, tail dependence 
asymmetry statistics
even had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a C copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population model outputs; and likewise for SG and SJ copulas. Dynamics were
strongly nonlinear for the model parameters we used, and nevertheless similarities
between noise and model-output copula structure, according to the statistics 
we used, were more noticable than discrepancies.
We again repeated the analysis with $r=1.3$. Discrepancies between noise and population
copula structure in this case were often glaring - apparently noise of standard deviation 1
interacted very strongly with model nonlinearities when the model was in its overcompensatory 
regime.

@@

### 2019-02-11 SupportingInformation.Rmd
changes to make ncsurrog throw an error when it should

<!--Basic setup-->
```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
seed<-101
# families<-c(1,3:10,13,14,16:20)
source("mtime.R") #A function needed for caching
```

@@

\newpage
```{r tab_count_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Lower-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged lower-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_LTdep_AICw}",
             booktabs=TRUE
             )
```

\newpage
```{r tab_count_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$UTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Upper-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged upper-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_UTdep_AICw}",
             booktabs=TRUE
             )
```

\newpage
```{r tab_count_LTmUT, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,
             digits = 3,linesep = "",
             format='latex',
             caption.short="Lower- minus upper-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_LTmUT}",
             booktabs=TRUE
             )
```

@@

\newpage
```{r tab_ff_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
             format='latex',linesep = "",digits=3,
             caption.short="AIC weights, N copulas, leaf-curling plum aphid first flight data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
             booktabs=TRUE
             )
```

@@

<!--
\newpage
## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}
### Model selection approach \label{Ex_pns_Model_selection}
### nonparametric analysis of tail dependence  \label{Ex_pns_NPA}
                          ----moved to Paper.Rmd----
-->

@@

  }
 }
}
```
<!--
\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
\caption[Comparing noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison). p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise cpoula family. \label{fig_Cause4copula_params_GOF_2500}}
\end{center}
\end{figure}
--->

@@

\newpage
```{r tab_temp_aphid_p_CvM, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$gfc_p_CvM,
             format='latex',
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**. Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_CvM}",
             booktabs=TRUE
             )
```

\newpage
```{r tab_temp_aphid_p_KS, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$gfc_p_KS,
             format='latex',
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**. Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_KS}",
             booktabs=TRUE
             )
```

\newpage
```{r tab_temp_aphid_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$LTdep_AICw,
             format='latex',
             caption = "Model-averaged lower tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTdep_AICw}",
             booktabs=TRUE
             )
```

\newpage
```{r tab_temp_aphid_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$UTdep_AICw,
             format='latex',
             caption = "Model-averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_UTdep_AICw}",
             booktabs=TRUE
             )
```

\newpage
```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw,
             digits = 5,
             format='latex',
             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTmUT}",
             booktabs=TRUE
             )
```
-->
<!--# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}-->

@@

- Given data $x_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$, let $\tau_{ij}$ be the Kendall
correlation of $x_i(t)$ and $x_j(t)$.  
- Calculate the covariance matrix of a multivariate normal distribution with standard-normal
marginals and with pairwise Kendall correlations equal to the $\tau_{ij}$. This is possible
using the `iTau` function of the `VineCopula` package.  
- Generate data $a_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$ by taking $T$ independent draws
from the the multivariate normal distribution of the previous step.  
- For any $i,t$ for which the datum $x_i(t)$ was missing, replace $a_i(t)$ by `NA`.
- For each $k$ and $i$ for which $a_i(t)$ is not `NA`, replace the 
$k$th-smallest (non-`NA`) element of the time series 
$a_i(t)$, $t=1,\ldots,T$ by the $k$th smallest (non-`NA`) 
element of $x_i(t)$, $t=1,\ldots,T$.
Call the result $b_i(t)$. This is the surrogate dataset.

Our code for this algorithm is `ncsurrog` in the `BIVAN` repository 
(\url{https://github.com/sghosh89/BIVAN}). Because, for each $i$, the final time series 
$b_i(t)$ is a permuted version of $x_i(t)$ (and with `NA`s for the same $t$), 
the marginal distributions of the surrogate dataset (i.e., the distributions of 
values for each sampling location, $i$) are exactly the same as for the original data.
Because ranks are preserved in passing from $a_i(t)$ to $b_i(t)$, pairwise Kendall 
correlations for the $b_i(t)$ will be the same as those of the $a_i(t)$. 
Discrepancies will come only from the presence of NAs and from sampling variation, 
and should be minor if the 
numbers of NAs are not too large and $T$ is not too small.

>>>>

<!--Basic setup-->
```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
seed<-101
# families<-c(1,3:10,13,14,16:20)
source("mtime.R") #A function needed for caching
library(captioner)
```

@@

<!--\newpage-->
```{r tab_count_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Lower-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged lower-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_LTdep_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_count_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$UTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Upper-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged upper-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_UTdep_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_count_LTmUT, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,
             digits = 3,linesep = "",
             format='latex',
             caption.short="Lower- minus upper-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_LTmUT}",
             booktabs=TRUE
             )
```

@@

<!--\newpage-->
```{r tab_ff_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
             format='latex',linesep = "",digits=3,
             caption.short="AIC weights, N copulas, leaf-curling plum aphid first flight data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
             booktabs=TRUE
             )
```

@@

@@

  }
 }
}
```
<!--
\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
\caption[Comparing noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison). p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise cpoula family. \label{fig_Cause4copula_params_GOF_2500}}
\end{center}
\end{figure}
-->

@@

<!--\newpage-->
```{r tab_temp_aphid_p_CvM, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$gfc_p_CvM,
             format='latex',
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**. Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_CvM}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_temp_aphid_p_KS, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$gfc_p_KS,
             format='latex',
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**. Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_KS}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_temp_aphid_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$LTdep_AICw,
             format='latex',
             caption = "Model-averaged lower tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTdep_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_temp_aphid_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$UTdep_AICw,
             format='latex',
             caption = "Model-averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_UTdep_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw,
             digits = 5,
             format='latex',
             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTmUT}",
             booktabs=TRUE
             )
```

<!--# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}-->

@@

- Given data $x_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$, let $\tau_{ij}$ be the Spearman (or Kendall) 
correlation of $x_i(t)$ and $x_j(t)$.
- Calculate the covariance matrix of a multivariate normal distribution with standard-normal
marginals and with pairwise Spearman (or Kendall) correlations equal to the $\tau_{ij}$. This is possible
using the `iRho` function (or the `iTau function) of the `VineCopula` package
- Generate data $a_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$ by taking $T$ independent draws
from the the multivariate normal distribution of the previous step.
- For any $i,t$ for which the datum $x_i(t)$ was missing, replace $a_i(t)$ by `NA`.
- For each $k$ and $i$ for which $a_i(t)$ is not `NA`, replace the 
$k$th-smallest (non-`NA`) element of the time series 
$a_i(t)$, $t=1,\ldots,T$ by the $k$th smallest (non-`NA`) 
element of $x_i(t)$, $t=1,\ldots,T$.
Call the result $b_i(t)$. This is the surrogate dataset.

\noindent Our code for this algorithm is `ncsurrog` in the `BIVAN` repository 
(\url{https://github.com/sghosh89/BIVAN}). Because, for each $i$, the final time series 
$b_i(t)$ is a permuted version of $x_i(t)$ (and with `NA`s for the same $t$), 
the marginal distributions of the surrogate dataset (i.e., the distributions of 
values for each sampling location, $i$) are exactly the same as for the original data.
Because ranks are preserved in passing from $a_i(t)$ to $b_i(t)$, pairwise Spearman 
(or Kendall) 
correlations for the $b_i(t)$ will be the same as those of the $a_i(t)$. 
Discrepancies between the $\tau_{ij}$ and the pairwise Spearman (or Kendall) 
correlations of the
$b_i(t)$ will therefore come about only from the presence of NAs and from sampling variation. 
These discrepancies should be minor if the 
numbers of NAs are not too large and $T$ is not too small. The Kendall version of the
algorithm throws an error if there are ties in any of the time series. The algorithm
also throws an error in the event that
the matrix generated in the second step is not positive semidefinite.

### 2019-02-11 makefile.Rmd
various changes
changes to make ncsurrog throw an error when it should

---
title: "Makefile_linux, compiliation notes"
author: "Daniel Reuman"
date: ""
output: pdf_document
---

@@

```{r tex_for_supp_mat}
rmarkdown::render(input="SupportingInformation.Rmd") 
```

>>>>

---
title: "BIVAN makefile, compiliation notes"
author: "Daniel Reuman"
date: ""
output: pdf_document
---

<!--checkpoint package-->
```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
library(checkpoint)
if (!dir.exists("./.checkpoint/")){
  dir.create("./.checkpoint/")
}
checkpoint("2018-03-29",checkpointLocation = "./")
```

@@

```{r tex_for_supp_mat}
library(rmarkdown)
rmarkdown::render(input="SupportingInformation.Rmd") 
```
